---
title: "Random Forest"
author: "Keen"
date: "5/18/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
suppressMessages(library(lattice))
suppressMessages(library(dplyr))
suppressMessages(library(tidyverse))
suppressMessages(library(ggplot2))
suppressMessages(library(MASS))
suppressMessages(library(tree))
suppressMessages(library(mosaic))
suppressMessages(library(randomForest))
library(rpart)
```

## Exploratory Data Analysis


Load the data. 
```{r}
#Lets take a look at the data
attrition = read.csv("WA_Fn-UseC_-HR-Employee-Attrition.csv")
#View(attrition)
```


Graph on the 
```{r}
#Check for missing values
#is.na(attrition)
ggplot(attrition, aes(x = MonthlyIncome)) + geom_histogram()
```
Histogram shows the distribution of monthly income in our dataset. The histogram is left skewed therefore we see that monthly income 

```{r}
# Age against Monthly income per education field
Age = attrition$Age
MonthlyIncome = attrition$MonthlyIncome
EducatioField = attrition$EducationField

qplot(Age, log(MonthlyIncome), data = attrition, color = EducationField)
```
The graphs above shows a positive relationship between age and monthly income across different kinds of educational fields.

```{r}
# Years since last promotion vs Years with current manager
qplot(YearsSinceLastPromotion, YearsWithCurrManager, data = attrition)
```
There does not seem to be any significant relationship between years since last promotion and years with current manager due to the random distribution shown in the plot

```{r}
#Attrition vs Number of companies worked

hase = ggplot(attrition, aes(x = Attrition, y = NumCompaniesWorked)) +
  geom_boxplot()
hase
```
The boxplot above shows that most employees lost to attrition have worked fewer companies than those who are not.

```{r}
#Attrition vs Distance from Home

hose = ggplot(attrition, aes(x = Attrition, y = DistanceFromHome)) + geom_boxplot()
hose
```
The graphs above suggests that the greater the distance from home the more likely it is for an employee to be lost through attrition.

```{r}
require(ggthemes)
ggplot(attrition,aes(DistanceFromHome,fill=Attrition))+
  geom_density(alpha=0.5)+
  theme_few()+
  theme(legend.position="bottom",plot.title=element_text(hjust=0.5,size=16))+
  labs(x="Distance from Home",title="Attrition and Distance from Home")
```

```{r}
#Attition vs Daily Rate/Distance From Home/ Houtly Rate
old.par = par()
par(mfrow = c(1,3))
with (attrition, {
  boxplot(attrition$DailyRate ~ attrition$Attrition, xlab = "Attrition by Daily Rate")
  boxplot(attrition$HourlyRate ~ attrition$Attrition, xlab = "Attrition by Hourly Rate")
  boxplot(attrition$MonthlyRate ~ attrition$Attrition, xlab = "Attrition by Monthly Rate")
}
)
par(old.par)
```

From the boxplots above, it seems attrition is most significantly affected by the daily rate as those who faced attrition had a significantly lower mean for daily rate as opposed to those who did not. As for attrition by hourly rate and by monthly rate, there does not seem to be any significant difference in means.

```{r}
#Work Life Balance vs Status
gaal = ggplot(attrition, aes(x = MaritalStatus, y = WorkLifeBalance)) + geom_boxplot()
gaal
```

The boxplot suggests that Work and Life Balance are irrespective of marital status as the means across the three different marital groups are the same

```{r}
#Job role vs percent salary hike
Haha = favstats(~attrition$PercentSalaryHike|attrition$JobRole)
Haha

bwplot(attrition$PercentSalaryHike ~ attrition$JobRole, data=attrition)

#attrition vs percent salary hike
Hehe = favstats(~attrition$PercentSalaryHike|attrition$Attrition)
Hehe
bwplot(attrition$PercentSalaryHike ~ attrition$Attrition, data=attrition)

#Gender by Montly Income
bwplot(attrition$MonthlyIncome ~ attrition$Gender, data=attrition)
```

Based on the tables above, Percent Salary Hike is independent of Job Role and is not a major factor in determining attrition as the means are almost the same between those employees lost to attrition and those not. It is also interesting from the Monthly income by gender boxplot that both men and women seem to be earning roughly the same amount which is usually not the case.

```{r}
Perfomance = attrition$PerformanceRating
MonthlyIncome = attrition$MonthlyIncome
Monthly.rate = attrition$MonthlyRate

qplot(Monthly.rate, MonthlyIncome, data = attrition, color = Perfomance)

old.par = par()
# create a grid of one row and two columns
par(mfrow = c(1,2))
with(c,{ggplot(c,aes(y = MonthlyIncome, x = log(Monthly.rate))) +
  geom_point()
  p1 +
  geom_point(aes(color = "Perfomance == 3"))
  ggplot(c,aes(y = MonthlyIncome, x = log(Monthly.rate))) +
  geom_point()
  p1 +
  geom_point(aes(color = "Perfomance == 4"))
}
)

c = data.frame(Monthly.rate, MonthlyIncome, Perfomance)
ggplot(c,
       aes(y = MonthlyIncome, x = log(Monthly.rate))) +
  geom_point()

p1 = ggplot(c, aes(x = log(Monthly.rate), y = MonthlyIncome))

p1 +
  geom_point(aes(color = Perfomance))

c$pred.SC = predict(lm(MonthlyIncome ~ log(Monthly.rate), data = c))

p1 + geom_point(aes(color = Perfomance)) +
  geom_line(aes(y = c$pred.SC))
```

From the above scatterplot, we can see that there is no linear relationship between monthly income and monthly rate

```{r}
counts = table(attrition$Attrition, attrition$BusinessTravel)
barplot(counts, main = "Attrition vs Business Travel", xlab="Number of attrition", col=c("darkblue","red"),
 	legend = rownames(counts))

S=favstats(~attrition$JobInvolvement|attrition$Attrition)
S

T=favstats(~attrition$JobSatisfaction|attrition$Attrition)
T
```


##Regressions

**Linear**

```{r}
names(attrition)
mod.lin <- lm(Age~ DailyRate+DistanceFromHome+JobRole+YearsAtCompany,data=attrition)
summary(mod.lin)
```


**Logistic**

```{r}
mod.log <- glm(Age+TotalWorkingYears+MonthlyRate~Attrition,data=attrition)
summary(mod.log)
```


**Simple plot using train data: Here I am interested in looking at how MonthlyIncome is affected by Age in terms of the TotalWorkingYears. Later, we are gunna try to predict the relationship of this and try to cluster. We can define the number of cluster that we want in terms of the TotalWorkingYears. We are gunna try to cluster this in a moment.**

```{r}
ggplot(attrition)+
  geom_point(aes(Age,MonthlyIncome,color=factor(TotalWorkingYears)))
```



##Let's just do a bit of data clean up. We can try to put this in four categories. 
```{r}
##put in data frame
attrition <- data.frame(attrition)
View(attrition)

attrition$Education <- as.factor(attrition$Education)
attrition$EnvironmentSatisfaction <- as.factor(attrition$EnvironmentSatisfaction)
attrition$JobInvolvement <- as.factor(attrition$JobInvolvement)
attrition$JobLevel <- as.factor(attrition$JobLevel)
attrition$JobSatisfaction <- as.factor(attrition$JobSatisfaction)
attrition$PerformanceRating <- as.factor(attrition$PerformanceRating)
attrition$RelationshipSatisfaction <- as.factor(attrition$RelationshipSatisfaction)
attrition$StockOptionLevel <- as.factor(attrition$StockOptionLevel)
attrition$WorkLifeBalance <- as.factor(attrition$WorkLifeBalance)

attrition$EmployeeCount <- NULL
attrition$EmployeeNumber <- NULL
attrition$StandardHours <- NULL
attrition$Over18 <- NULL

```

Split data into train and test train.
```{r}
n0 <- nrow(attrition)
train <- sample(1:n0,n0/2,rep=F)
train.df <- attrition[train,]
test.df <- attrition[-train,]
#names(train.df)
```

Decision Tree

Here, we are gunna predict Atttrition using decision tree and see which variable is the most important in predicting whether or nott someone will leave their job. 

```{r}
mod.tree <- tree(Attrition~., data=train.df,control=tree.control(nrow(attrition),mindev = 0.01))
summary(mod.tree)
```


Let's see the tree
```{r}
plot(mod.tree)
text(mod.tree, pretty=0,cex=0.5)
```
Observation: Seems like some of the most important variable for Attrition are YearsAtCompany, PercentSalaryHike, YearsSinceLastPromotion, MonthlyRate. 

Predict train data. how well was our train data?
```{r}
##make some predictions with test data
preds.train<- predict(mod.tree,type="class")

##confusion matrix
with(train.df,table(Attrition,preds.train))

(err.train<- with(train.df,mean(Attrition!=preds.train)))
```

Conclusion: Our train data was pretty good with error rate of 0.09931973

How about wih our test data?
```{r}
##make some predictions with test data
preds.test<- predict(mod.tree,newdata=test.df,type="class")
##confusion matrix
with(test.df,table(Attrition,preds.test))

(err.test <- with(test.df,mean(Attrition!=preds.test)))
```
Conclusion: Not bad, I think maybe we can do better?


Here, we are gunna try with smaller mindev, how does err.test change?
```{r}
mod.tree2 <- tree(Attrition~.,data=train.df,
                  control=tree.control(nrow(train.df),mindev=0.0000000))
(numLeaves <- sum(mod.tree2$frame$var =="<leaf>"))

<<<<<<< HEAD
numFolds <- 10
=======
numFolds <- 25
>>>>>>> 3e2b84e22aa3f575ea76cd2643548a0ee0591fa1
folds <- sample(1:numFolds,nrow(train.df),rep=T)

err <- matrix(nrow=numLeaves,ncol=3)
errCV <- numeric(numFolds)
for(treeSize in numLeaves:2){
    print(treeSize)
    model.prune <- prune.tree(mod.tree2,best=treeSize)
    preds.train <- predict(model.prune,newdata=train.df,type="class")
    preds.test <- predict(model.prune,newdata=test.df,type="class")
    errTrain <- with(train.df,mean(Attrition!=preds.train))
<<<<<<< HEAD
    errTest <- with(test.df,mean(Attrition!=preds.test))
    ##Cross validate: 
    
=======
    errTest <- with(test.df,mean(Attrition !=preds.test))
    ##Cross validate: 
>>>>>>> 3e2b84e22aa3f575ea76cd2643548a0ee0591fa1
    for(fold in 1:numFolds){
        attTrainTrain.df <- train.df[fold != folds,]
        attTrainTest.df <- train.df[fold == folds,]
        att.cv <- tree(Attrition~.,data=attTrainTrain.df,
                          control=tree.control(nrow(attTrainTrain.df),mindev=0.0000000))
        att.cv.prune <- prune.tree(att.cv,best=treeSize)
        preds <- predict(att.cv.prune,newdata=attTrainTest.df,type="class")
        errCV[fold] <- with(attTrainTest.df,mean(Attrition!= preds))
    }
    err[treeSize,] <- c(errTrain,errTest,mean(errCV))
}

data.frame(treeSize=2:numLeaves,
           train=err[-1,1],
           test=err[-1,2],           
           cv=err[-1,3]) %>%
    gather(type,val,train:cv) %>% 
    ggplot()+
    geom_point(aes(treeSize,val,color=type),size=1)+
    geom_line(aes(treeSize,val,color=type))+    
    scale_color_manual(values=c("red","blue","black"))

```
Seems like test error did not change

Let's prune our tree! Single Cross validation for optimal size
```{r}
mod.cv <- cv.tree(mod.tree,FUN=prune.misclass)
plot(mod.cv$size,mod.cv$dev)
plot(mod.cv)
```

What is our optimal tree size?
```{r}
(opt.size <- mod.cv$size[which(mod.cv$dev==min(mod.cv$dev))])
size <- rev(mod.cv$size)
(opt.size2 <- size[which.min(rev(mod.cv$dev))])
```

```{r}
par(mfrow=c(1,2))
plot(mod.cv$size,mod.cv$dev,type="b")
plot(mod.cv$k,mod.cv$dev,type="b")

```

Now let us prune this using the optimal size
```{r}
mod.prune <- prune.misclass(mod.tree,best=opt.size)
plot(mod.prune)
text(mod.prune,pretty=0,cex=0.5)

```
Observation: 

Seems like the important variable for employees staying at their company is HourlyRate, RelationshipSatisfacion, YearsAtCompany, and EnvironmentSatisfication. 




Predict our prune model using the testing data. What do we find?
```{r}
mod.pred.prune <- predict(mod.prune,test.df,type="class")
summary(mod.pred.prune)

##confusion matrix
with(test.df,table(Attrition,mod.pred.prune))

(err.prune.test<- with(test.df,mean(Attrition!=mod.pred.prune)))
```


##compare error rates from the tree and the pruned tree

```{r}
c(err.train,err.test,err.prune.test)
```
Looks like we did a better job finding the most important variable for predicting attrition when we pruned our tree and found the best size. 


Let's predict our prune tree with out testing data set. How does it look?
```{r}
probs.prune <- predict(mod.prune,newdata=test.df)
hist(probs.prune[,1],breaks=25)
```


##Random Forest 

Here we are going to perform some out of bag error. What does random forest say the most important variable for predicting Attriion. 
```{r}
p <- ncol(train.df)-1
mod.rf <- randomForest(Attrition~., data = train.df, ntree = 500, mtry = p,importance = TRUE)
mod.rf

```


Now, let's make a prediction using our testing data and see how well we did. 
```{r}
## Make some predictions on the test data
preds.bag <- predict(mod.rf,newdata=test.df,ntree=500)

## Confusion matrix
with(test.df,table(Attrition,preds.bag))

(err.bag <- with(test.df,mean(Attrition!=preds.bag)))

```


Let's compare the error rates from our decision tree, prune tree, and the out of bag error
```{r}
c(err.test,err.prune.test,err.bag)
```



```{r}
plot(mod.rf)
```

```{r}
importance(mod.rf)
```

```{r}
varImpPlot(mod.rf)
```

```{r}
require(dplyr)
attrition <- attrition %>%
      mutate(Attrition = ifelse(Attrition== "No",0,1))
View(attrition)
```

```{r}
library(gbm)
numTrees <- 100
depth <- 7
lambda <- .1
att.gbm <- gbm(Attrition ~ ., data=attrition,
                n.trees=numTrees,
                distribution="adaboost",
                interaction.depth = depth,
                shrinkage=lambda)

```


```{r}
probs <- predict(att.gbm,newdata=attrition,n.trees=100,type="response")
##convert to a prediction.
preds <-  probs > 0.5

## How does it look?
with(attrition,table(Attrition,preds))
with(attrition,mean(Attrition != preds))
```

 
A quick look at the error as a function of the number of trees.
```{r}
errTree <- numeric(numTrees)
for(n in 1:numTrees){
  probs <- predict(att.gbm,newdata=attrition,n.trees=n,type="response")
  preds <-  probs > 0.5
  errTree[n] <- with(attrition,mean(Attrition != preds))  
}


data.frame(trees=1:numTrees,err=errTree) %>% 
  ggplot()+
  geom_point(aes(trees,err))

```

```{r}
summary(att.gbm)
```




