---
title: "TariAnalysis"
output: html_document
editor_options: 
  chunk_output_type: console
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggplot2)
library(lattice)
library(mosaic)
library(MASS)
library(tree)
library(randomForest)
library(dplyr)
library(ggrepel)
library(graphics)
library(factoextra)
```

## Exploratory Data Analysis

```{r}
#Lets take a look at the data
attrition = read_csv("WA_Fn-UseC_-HR-Employee-Attrition.csv")
head(attrition)
summary(attrition)
str(attrition)
```

```{r}
#Check for missing values
is.na(attrition)
ggplot(attrition, aes(x = MonthlyIncome)) + geom_histogram()
```
Histogram shows the distribution of monthly income in our dataset. The histogram is left skewed therefore we see that monthly income 

```{r}
# Age against Monthly income per education field

Age = attrition$Age
MonthlyIncome = attrition$MonthlyIncome
field = attrition$EducationField

qplot(Age, log(MonthlyIncome), data = attrition, color = EducationField)
```
The graphs above shows a positive relationship between age and monthly income across different kinds of educational fields.

```{r}
# Years since last promotion vs Years with current manager
qplot(YearsSinceLastPromotion, YearsWithCurrManager, data = attrition)
```
There does not seem to be any significant relationship between years since last promotion and years with current manager due to the random distribution shown in the plot

```{r}
#Attrition vs Number of companies worked

hase = ggplot(attrition, aes(x = Attrition, y = NumCompaniesWorked)) +
  geom_boxplot()
hase
```
The boxplot above shows that most employees lost to attrition have worked fewer companies than those who are not.

```{r}
#Attrition vs Distance from Home

hose = ggplot(attrition, aes(x = Attrition, y = DistanceFromHome)) + geom_boxplot()
hose
```
The graphs above suggests that the greater the distance from home the more likely it is for an employee to be lost through attrition.

```{r}
#Attition vs Daily Rate/Distance From Home/ Houtly Rate
old.par = par()
par(mfrow = c(1,3))
with (attrition, {
  boxplot(attrition$DailyRate ~ attrition$Attrition, xlab = "Attrition by Daily Rate")
  boxplot(attrition$HourlyRate ~ attrition$Attrition, xlab = "Attrition by Hourly Rate")
  boxplot(attrition$MonthlyRate ~ attrition$Attrition, xlab = "Attrition by Monthly Rate")
}
)
par(old.par)
```

From the boxplots above, it seems attrition is most significantly affected by the daily rate as those who faced attrition had a significantly lower mean for daily rate as opposed to those who did not. As for attrition by hourly rate and by monthly rate, there does not seem to be any significant difference in means.

```{r}
#Work Life Balance vs Status

gaal = ggplot(attrition, aes(x = MaritalStatus, y = WorkLifeBalance)) + geom_boxplot()
gaal
```

The boxplot suggests that Work and Life Balance are irrespective of marital status as the means across the three different marital groups are the same

```{r}
#Job role vs percent salary hike
Haha = favstats(~attrition$PercentSalaryHike|attrition$JobRole)
Haha

bwplot(attrition$PercentSalaryHike ~ attrition$JobRole, data=attrition)

#attrition vs percent salary hike
Hehe = favstats(~attrition$PercentSalaryHike|attrition$Attrition)
Hehe
bwplot(attrition$PercentSalaryHike ~ attrition$Attrition, data=attrition)

#Gender by Montly Income
bwplot(attrition$MonthlyIncome ~ attrition$Gender, data=attrition)
```

Based on the tables above, Percent Salary Hike is independent of Job Role and is not a major factor in determining attrition as the means are almost the same between those employees lost to attrition and those not. It is also interesting from the Monthly income by gender boxplot that both men and women seem to be earning roughly the same amount which is usually not the case.

```{r}
Perfomance = attrition$PerformanceRating
MonthlyIncome = attrition$MonthlyIncome
Monthly.rate = attrition$MonthlyRate

qplot(Monthly.rate, MonthlyIncome, data = attrition, color = Perfomance)

old.par = par()
# create a grid of one row and two columns
par(mfrow = c(1,2))
with(c, {
      ggplot(c,
       aes(y = MonthlyIncome, x = log(Monthly.rate))) +
  geom_point()
  p1 +
  geom_point(aes(color = "Perfomance == 3"))
  ggplot(c,
       aes(y = MonthlyIncome, x = log(Monthly.rate))) +
  geom_point()
  p1 +
  geom_point(aes(color = "Perfomance == 4"))
}
)

c = data.frame(Monthly.rate, MonthlyIncome, Perfomance)
ggplot(c,
       aes(y = MonthlyIncome, x = log(Monthly.rate))) +
  geom_point()

p1 = ggplot(c, aes(x = log(Monthly.rate), y = MonthlyIncome))

p1 +
  geom_point(aes(color = Perfomance))

c$pred.SC = predict(lm(MonthlyIncome ~ log(Monthly.rate), data = c))

p1 + geom_point(aes(color = Perfomance)) +
  geom_line(aes(y = c$pred.SC))
```

From the above scatterplot, we can see that there is no linear relationship between monthly income and monthly rate

```{r}
counts = table(attrition$Attrition, attrition$BusinessTravel)
barplot(counts, main = "Attrition vs Business Travel", xlab="Number of attrition", col=c("darkblue","red"),
 	legend = rownames(counts))

S=favstats(~attrition$JobInvolvement|attrition$Attrition)
S

T=favstats(~attrition$JobSatisfaction|attrition$Attrition)
T
```


##Regressions

**Linear**

```{r}
names(attrition)
mod.lin <- lm(Age~ DailyRate+DistanceFromHome+JobRole+YearsAtCompany,data=attrition)
summary(mod.lin)
```


**Logistic**

```{r}
mod.log <- glm(Age~ Attrition,data=attrition)
summary(mod.log)
```


**KNN**

```{r}

```


##Classifications

**KNN**
```{r}


```



**LDA**
```{r}


```

##Cluster

##Random Forest

##SVM, PCA, SVD, ADABoosting

##PCA

Before trying to cluster since there are a lot of variables lets try to reduce the number of variables through picking the most important ones

change order of columns
```{r}
attri2 <- attrition[,c(2:34,1)]
attri2 <- attri2[,-c(21)]
```


```{r}
sampleSize <- nrow(attrition)
```


```{r}
mseBoot <- function(data.df,M=50){
    mse <- rep(0,M)
    for(m in 1:M){
        bootSamp <- sample(1:sampleSize,sampleSize,rep=T)
        outOfBag <- setdiff(1:sampleSize,bootSamp)
        train.df <- data.df[bootSamp,]
        test.df <-   data.df[outOfBag,]
        mod.rf <- randomForest(Age~.,data=train.df,ntree = 500, mtry = p,importance = TRUE)
        vals <- predict(mod.rf,newdata=test.df, ntree=500)
        mse[m] <- with(test.df,mean((Age-vals)^2))
    }
    mean(mse)
}
```


```{r}
numPreds <- length(names(attri2))-1
```

```{r}
allPreds <- 1:(numPreds)
```

```{r}
currPreds <- c()
```

```{r}
availPreds <- setdiff(allPreds,currPreds)
length(availPreds)
```

```{r}
maxPreds <- numPreds
```

```{r}
minMSE <- numeric(maxPreds)
```

```{r}
tot <- 0
while( tot < maxPreds){
    ##add predictor which decreases MSE (as determined by CV or
    ##Bootstrapping)
    ## The MSEs computed as we add each of the available predictors
    allMSE <- numeric(length(availPreds))
    ctr<-1
    for(id in availPreds){
        data.df <- attri2[,c(currPreds,id,numPreds+1)]
        mse <- mseBoot(data.df,30)
        allMSE[ctr] <- mse
        ctr<-ctr+1
    }
    ##Find the min
    id <- which.min(allMSE)
    ##get the best predictor and MSW
    bestPred <- availPreds[id]
    bestMSE <- min(allMSE)
    ##Add these into the collection
    currPreds <- c(currPreds,bestPred)
    tot <-tot+1
    minMSE[tot] <- bestMSE
    availPreds <- setdiff(allPreds,currPreds)
    ## Print stuff out for debugging and attention-grabbing
    print(sprintf("Predictor Added: %s  MSE Value: %s",bestPred,bestMSE))
    print(currPreds)
}
```
 It works, so we can now look at MSE Plot
 
```{r}
Titles <- names(attri2[,c(currPreds)])
```
 
change in MSE
```{r}
diffMSE <- c(0,minMSE[-length(minMSE)]-minMSE[-1])
head(diffMSE)
```

```{r}
result.df <- data.frame(id=1:length(Titles),Titles,minMSE,diffMSE=-10*diffMSE, predictors= currPreds)
```

```{r}
cut <- 10
ggplot(result.df %>%
       filter(Titles %in% Titles[1:cut]))+
    geom_segment(aes(x=1,xend=id,y=0,yend=0),size=.1,color="black")+
    geom_segment(aes(x=id,xend=id,y=0,yend=minMSE),color="blue")+
    geom_point(aes(x=id,y=minMSE),color="blue",size=1.5)+
    geom_segment(aes(x=id,xend=id,y=0,yend=diffMSE),color="red")+
    geom_point(aes(x=id,y=diffMSE),color="red",size=1.5)+
    scale_x_continuous(breaks=1:length(Titles),label=Titles)+
    coord_flip()+
    ggtitle("Subset Selection: MSE Decrease")
```

These are the top 10 variables that can be used to predict someone's age based on attrition dataset. The bluw part shows the overall reduction in MSE whilst the red part shows the difference in MSE with the use of each variable.

We can now use those variables in our PCA to see how different ages cluster together

```{r}
result.df <- result.df %>%
  filter(minMSE <= quantile(result.df$minMSE, 0.25))
```



```{r}
attri.df <- attrition[,result.df$predictors]
ages <- attrition$Age
attri.df <- data.frame(select_if(attri.df, is.numeric))
```

```{r}
attri.pca <- prcomp(attri.df)
fviz_pca_biplot(attri.pca,repel=TRUE)
```
This is really intense and hard to understand, it seems like most of the variables fit into one dimension, I'll try reducing the amount to see what we get.

Pull off the PCA information
This is the change of basis matrix (V in the SVD)
```{r}
rots <- attri.pca$rotation
```


Rotate into the new basis.
```{r}
attri.pca1<- data.matrix(attri.df) %*% rots
dim(attri.pca1)


```


The first three directions
```{r}
pca1 <- attri.pca1[,1]
pca2 <- attri.pca1[,2]
pca3 <- attri.pca1[,3]
```


##Add in K Means clustering to the Principal Components


2 clusters, this is arbitrary
```{r}
K <- 2
```


run 30 times, allow up to 30  iterations for convergence
```{r}
attri.km <- kmeans(attri.df,K,iter.max=20,nstart=30)

```


Our clusters
```{r}
attri.km$cluster

```


Add the clusters to the data frame
```{r}
attri.df$cluster.km<- attri.km$cluster
head(attri.df)

```


Look at one cluster
What ages land where?
```{r}

attri.df$age <- attrition$Age
filter(attri.df,cluster.km==1)%>%
    dplyr::select(age)


```


What are the cluster sizes?
```{r}
with(attri.df,table(cluster.km))

```


Build the rotated data frame
```{r}
attriCluster.df <- data.frame(pca1,pca2,cluster=factor(attri.df$cluster.km),name=ages)

```


plot the cluster.pcas on top of the PCA first two components
```{r}
ggplot(sample(attriCluster.df, (nrow(attriCluster.df)/4), replace = FALSE),aes(pca1,pca2,color=cluster))+
    geom_point(size=1)+
    geom_text_repel(aes(label=name),size=3)+
    ggtitle("Attrition Cluster Analsys and PCA")
```


```{r}
with(attri.df,table(ages, cluster.km))
```

Lets find the optimum number of clusters 
```{r}
fviz_nbclust(attri.df,kmeans,method="wss")
```

From the graph it looks like the optimum is 5 so we will cluster at 5 and see what that gives us

```{r}
k <- 5
attri.km <- kmeans(attri.df,K,iter.max=20,nstart=30)
attri.km$cluster
attri.df$cluster.km<- attri.km$cluster
with(attri.df,table(cluster.km))
attriCluster.df <- data.frame(pca1,pca2,cluster=factor(attri.df$cluster.km),Age=ages)
ggplot(attriCluster.df,aes(pca1,pca2,color=cluster))+
    geom_point(size=1)+
    geom_text_repel(aes(label=name),size=3)+
    ggtitle("Attrition Cluster Analsys and PCA")
```

From the clusters it looks like age is not a very good variable for differentiating employees as there is a pretty even spread of ages between each cluster


The table below better exemplifies this
```{r}
with(attriCluster.df,table(Age, cluster))
```
So we can conclude form the data, in this case Age itself does not seperate employee groups, as you can be similar to any other employee regardless of your age.
